Course 5: Reproducible Research
Week 2 Project

Download and unzip data file code
```{r downloaddata}
url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url,destfile="./Data/Dataset.zip",sep="/")
unzip(zipfile="./data/Dataset.zip",exdir="./data")

data <-read.csv("./data/activity.csv")
```

2. Histogram of Daily Steps
```{r histogram}
library(dplyr)
data2 <- data %>% select(date, steps) %>% group_by(date) %>% summarize(total_steps = sum(steps)) %>% na.omit()
hist(data2$total_steps, main = "Histogram of Total Daily Steps", col =4, xlab = "Steps", breaks=20)
```

3. Mean and median number of steps taken each day in sample
```{r meanmedian}
mean(data2$total_steps)
median(data2$total_steps)
```

4. Time series plot of average number of steps taken
```{r timeseries}
data3 <- data %>% select(interval, steps) %>% na.omit() %>% group_by(interval) %>% summarize(ave_steps = mean(steps))

library(ggplot2)
ggplot(data3, aes(x=interval, y=ave_steps)) + geom_line() + 
  labs(y="Ave number of steps") +
  labs(title = "Average number of steps taken per day")
```
5. 5 minute interval that on average contains maximum number of steps
```{r max_interval}
subset(data3, data3$ave_steps==max(data3$ave_steps))
```

6. Imputing missing data
```{r missing}
missing <- sum(is.na(data$steps))
missing
```

2304 missing step counts. Strategy for imputation is to take the average for that 5 minute interval over the whole dataset and replace NAs and check no NAs remaining
```{r imputation}
data_mean <- data %>% group_by(interval) %>% 
  mutate(steps = ifelse(is.na(steps), mean(steps, na.rm =T), steps))
head(data_mean)
missing2 <- sum(is.na(data_mean$steps))
missing2
```

7. Histogram of steps taken after imputation
```{r hist after imp}
data5 <- data_mean %>% select(date, steps) %>% group_by(date) %>% summarize(total_steps = sum(steps))
hist(data5$total_steps, main = "Histogram of Total Daily Steps (after Imp)", col =4, xlab = "Steps", breaks=20)
```

Comparison of mean and median before and after imputation
```{r mean comparison}
mean_before <- mean(data$steps, na.rm = T)
mean_after <- mean(data_mean$steps)
median_before <- median(data$steps, na.rm = T)
median_after <- median(data_mean$steps)

mean_before
mean_after
median_before
median_after
```
No difference in mean or median, so it is a suitable imputation

8.  Panel plot of average number of steps at weekend and weekdays
```{r weekend weekday comparison}
data_mean$date <- as.Date(data_mean$date)
data_mean$weekday <- weekdays(data_mean$date)
data_mean$weekend <- ifelse(data_mean$weekday == "Saturday" | data_mean$weekday == "Sunday", "weekend", "weekday")

data_weekday <- data_mean %>% select(weekend, steps, interval) %>% group_by(weekend, interval) %>% summarize(ave_steps = mean(steps))

ggplot(data_weekday) + aes(x=interval, y=ave_steps, color = weekend) +
  geom_line() +
  facet_grid(weekend~.) +
  labs(title = "Average number of steps by weekend and weekday") +
  xlab("interval") +
  ylab("Average number of steps")
```
